### 2019-02-25 Writeup/PERT_Paper_Draft1.tex
Updated writeup + upload python deterministic tests

Optimizing Crashing Decisions in a Project Management Problem with Disruptions \ms

@@

	The management of complex projects through optimization has a rich history in operations research, beginning with the critical path method of \cite{kelley1961criticalpath} for the deterministic project management problems; see S{\"o}derlund \cite{soderlund2004building} for an overview. A project is a collection of activities \(I\) of some duration \(D_i, i \in I\), and there are precedence relationships, \(\mathcal{A} \subseteq I \times I\), between activities due to logical or technological considerations. A project can be characterized as an acyclic activity network \(\mathcal{G} = (I,\mathcal{A})\). An arc \((i,j) \in \mathcal{A}\) indicates that activity \(i\) has to be finished before the start of activity \(j\), and its length shows the duration of an activity \(D_i\). An activity cannot start until all its predecessors are completed, while multiple activities can be processed at the same time without a upper limit, as long as the precedence requirement is satisfied. More details about the activity network are discussed in Elmaghraby~\cite{Elmaghraby77}. In this setting, ``crashing" is an action that consumes certain amount of resource and shorten the duration of activities accordingly~\cite{kuhl2008dynamic}. We can apply a finite set of crashing options, \(j \in J_i\), to activity \(i \in I\). One unit application of each option incurs a cost of \(b_{ij}\), while it decreases the length of activity \(i\) by \(D_ie_{ij}\); i.e., the crashed duration is proportional to both the nominal activity duration and the crashing decision. The total cost of crashing options cannot exceed a given budget, \(B\). The deterministic optimization problem for crashing decisions has been proposed since the 1960s \cite{fulkerson1961network, kelley1961criticalpath}. The goal is to complete all activities in minimum time by allocating resources under budget constraints. We can create two dummy activities \(S\) and \(T\) to represent the start and the end of entire project. \(S\) should precede every activity \(i \in I \backslash \{S\}\) and \(T\) should succeed every activity \(i \in I \backslash \{T\} \). Then the objective is then equivalent to minimize the starting time of activity \(T\). Thus, we formulate the deterministic project crashing optimization problem as:

@@

		\text{s.t.} \quad &  t_k - t_i \geq D_{i}(1 - \sum_{j \in J_i} x_{ij} e_{ij}) \qquad \qquad \forall \,i \in I, (i,k) \in \mathcal{A} \label{cons:dSep}\\

@@

		& t_i \geq 0  \qquad \qquad \forall \,i \in I \label{cons:dNonnegt}\\
		& 0 \leq x_{ij} \leq 1  \qquad \qquad \forall \,i \in I, j \in J_i. \label{cons:dxub}

@@

	In this formulation, \(t_i\) represents the starting time of activity \(i \in I\). We aim to minimize the project span, which is the starting time of the terminal activity, \(t_T\). Constraint~\eqref{cons:dSep} guarantees the precedence relationship: if activity \(i\) precedes activity \(k\), activity \(k\) cannot start until activity \(i\) is finished, since the finishing time of activity \(i\) is \(t_i + D_i (1- \sum_{j \in J_i} x_{ij}e_{ij})\). Constraint~\eqref{cons:dBudget} is the budget constraint and constraint~\eqref{cons:dSingleBudget} ensures that no more than one unit of crashing option can be applied to an activity. In constraint~\eqref{cons:dxub}, we can put an artificial upper bound \(1\) on all crashing decisions \(x_{ij}\) by taking proper values of \(e_{ij}\). \\

@@

	The deterministic model can be extended to incorporate uncertainty of activity durations, starting with the program evaluation and review technique (PERT) \cite{malcolm1959application}. The PERT model assumes that activities are independently distributed with beta distributions, and the expected project span can be approximated by a normal distribution. However, this assumption may not be accurate for a general case~\cite{Elmaghraby77}. Therefore, assuming the activity durations follow a given probability distribution, Monte Carlo simulation methods can be used to estimate the expected project span, since it is is difficult to express analytically the expected project span \cite{burt1971conditional,van1963letter}. Heuristics and simulation-based algorithms have been developed to solve the stochastic project crashing problem \cite{aghaie2009ant, bowman1994stochastic, ke2014genetic, kim2007heuristic}. Another approach to handle the uncertainty is robust optimization, in which the objective is to minimize the worst case project span within a specified uncertainty set. While affinely adaptive recourse decisions are computationally tractable as linear, or second-order cone, programs, this restriction may lead to suboptimal solutions \cite{chen2008linear, cohen2007stochastic}. However, once recourse decisions can take general form, the robust model is only tractable with hypercube uncertainty sets, which is a trivial case because the worst-case scenario in the uncertainty set is always the point where every activity takes its upper bound \cite{wiesemann2012robust}. Ahipasaoglu et al. \cite{ahipasaoglu2016distributionally} propose a distributionally robust optimization scheme applied to a PERT network, which reformulates the problem as a semidefinite program or a copositive program, depending on the description of uncertainty. The project crashing optimization problem finds wide applications in project management~\cite{demeulemeester2006project, jaselskis1991allocation,tonchia2018industrial}, machine scheduling~\cite{blazewicz1983scheduling,hall1996machine}, health services scheduling~\cite{cardoen2010operating}, chemical processes~\cite{li2008process}, and digital circuit sizing~\cite{kim2007heuristic}.\\

@@

	In this paper, we propose using the concept of stochastic disruptions to model the uncertainty, which is different from the setting of a stochastic program or a robust optimization problem. A stochastic disruption is an event that may occur any time in the time horizon and change the system parameters significantly. A few authors apply this idea when the disruption can only occur in a set of specified time periods. Yu and Qi \cite{yu2004disruptionmgt} introduce scenario-based optimization models for airline scheduling. Morton et al. \cite{morton2009sealift} introduce a sealift scheduling problem under a finite number of stochastic disruptions within a stochastic programming structure. This structure ``falls between standard two-stage and multi-stage stochastic programs for a multi-period problem" and reduces the size of the problem to a quadratic growth in the number of time periods. Our setting inherits the philosophy of \cite{morton2009sealift} but enhances the model by allowing a continuous disruption time instead of a prespecified set of fixed time periods. \\

@@

	We first formally describe the crashing optimization problem under stochastic disruptions. Given a limited number of disruption scenarios, the problem can be formulated as a stochastic mixed integer program and we present the extensive formulation in Section~\ref{sec:formulation}. If we assume a continuous distribution for the disruption time and magnitude, a sample average approximation (SAA) can be used to create a finite set of scenarios and approximate the original problem by a finite-sized optimization problem. We present some important properties of this problem using a serial activity network as a special case in Section~\ref{sec:examples}. The large scale and the discrete non-convex nature of the SAA problem's extended formulation means it may not be solved efficiently to desirable tolerance level. In Section~\ref{sec:decomposition}, a branch-and-cut method based on Benders decomposition is developed to solve the crashing optimization problem with a disruption. We show such a decomposition method solves the integer program to the exactness within finite number of iterations. The experiment results are presented in Section~\ref{sec:results}, including the relationship between the solution quality and the sample size, the comparison between the quality of our solution and the deterministic/semi-stochastic alternative solutions, and the computational performance of the decomposition method presented in Section~\ref{sec:decomposition}. We conclude our paper with remarks on potential extensions of this model in Section~\ref{sec:conclusions}.

@@

	For a project crashing problem under stochastic disruptions, we assume at most one stochastic disruption can occur at a random time in the project span, since for the applications we are interested in (natural disasters, terrorist attacks, major market crashes, etc.), it is unlikely for two disruptions to occur in entire the time horizon. For example, suppose we manage a construction project, and we aim to plan against the potential hazard caused by earthquakes. It is unlikely for two major earthquakes to hit the same area twice within a year (aftershocks can be considered part of a major earthquake.) It is also reasonable to assume for each activity \(i \in I\), the crashing decision needs to be made at the start of that activity, and cannot be changed afterwards, especially when contracts are involved in commitment of resources~\cite{oberlender1993project}. The disruption does not affect the started activities but changes the length of those that have not started according to a random distribution. It is usually hard to compute the recourse function directly when parameters are distributed according to continuous distributions, and therefore SAA is proposed to approximate the problem by a large-scale deterministic optimization problem with a finite scenario set \cite{kim2015guide,shapiro2009lectures}. In this paper, we assume there is a discrete finite set of scenarios indexed by \(\omega \in \Omega\). For each scenario \(\omega\), the random realization of parameters, \(\xi^\omega\), consists of the timing of disruption, \(H\), and the changed duration of each activity, \(d_i, \forall i \in I\). \\

@@

	If there is at most one disruption, we can model the problem as a two-stage stochastic mixed integer program. The definition of our first stage is different from the traditional stochastic program setting. Here the first stage contains decisions through completion of the project if no disruption occurs and the second stage characterizes the decisions for each realization of the disruption. This means that the first stage decision variables are carried out until the disruption, if it ever occurs, and after the disruption, the scenario-specific recourse decisions are executed.\\

@@

		\(t_{i}\) & \(\qquad\) & nominal starting time of activity \(i\), \(i \in I\);&\\

@@

		\(t_{i}^\omega\) & \(\qquad\) & starting time of activity \(i\) under scenario \(\omega\), \(i \in I, \omega \in \Omega\);&\\

@@

@@

	In this section we show a serial example to illustrate the nature of the project crashing problem under stochastic disruption, which is different from the deterministic problem. In the deterministic project crashing problem, all activities on the critical path should start as soon as possible. However, with a stochastic disruption, it might be optimal to delay the start of some activities. Therefore, it is necessary to have decision variable \(t_i, \forall i \in I\) in model~\eqref{prob:extensive} in addition to the crashing decision variable \(x_{ij}, \forall i \in I, j \in J_i\). \\

@@

	Because it is possible for an optimal crashing plan to contain a delay for some activities, we have to set up decision variables \(t_i,\ \forall i \in I\), as the starting time of each activity, rather than assume that each activity starts as soon as all of its predecessors are finished.  We summarize the reasons of such delay include:

@@

			\item Some activities might have a shorter length after disruption. It is beneficial to wait a short period of time to capture this advantage.

>>>>

Optimal Crashing of an Activity Network with Disruptions \ms

@@

	The management of complex projects through optimization has a rich history in operations research, beginning with the critical path method of \cite{kelley1961criticalpath} (see S{\"o}derlund \cite{soderlund2004building} for an overview). A project is a collection of activities \(I\)
	 %of some duration \(D_i, i \in I\)
	 , and there are precedence relationships, \(\mathcal{A} \subseteq I \times I\), between activities due to logical or technological considerations. A project can be characterized by an acyclic activity network \(\mathcal{G} = (I,\mathcal{A})\). An arc \((i,k) \in \mathcal{A}\) indicates that activity \(i\) has to be finished before the start of activity \(k\), and its length shows the required separation between their start time,
	 %duration of an activity 
	 \(D_{ik}\). An activity cannot start until all its predecessors are completed, while multiple activities can be processed at the same time without an upper limit, as long as the precedence requirements are satisfied. See Elmaghraby~\cite{Elmaghraby77} for a detailed treatment of activity networks. In this setting, ``crashing" is an action that consumes a certain amount of resource and shortens the duration of activities accordingly~\cite{kuhl2008dynamic}. We can apply a finite set of crashing options, \(j \in J_i\), to activity \(i \in I\). One unit application of each option incurs a cost of \(b_{ij}\), \textcolor{blue}{while it decreases all required separation lengths starting from activity \(i\) by \(D_{ik}e_{ij},\ \forall (i,k) \in \mathcal{A}\); i.e., the crashed duration is proportional to both the nominal separation length and the crashing decision.} The total cost of crashing cannot exceed a given budget, \(B\). The deterministic optimization problem for crashing an activity network was proposed in the 1960s \cite{fulkerson1961network, kelley1961criticalpath}. The goal is to complete all activities in minimum time by allocating resources under one or more budget constraints. We create two dummy activities \(S\) and \(T\) to represent the start and the end of the entire project. Activity \(S\) should precede every activity \(i \in I \backslash \{S\}\) and \(T\) should succeed every activity \(i \in I \backslash \{T\} \), and they both have zero duration, \(D_{Si} = D_{iT} = 0,\ \forall i \in I\). Given $t_S \equiv 0$, the objective is then to minimize the start time of activity \(T\). Thus, we formulate the deterministic project crashing problem as:

@@

		\text{s.t.} \quad &  t_k - t_i \geq D_{ik}\left(1 - \sum_{j \in J_i} x_{ij} e_{ij}\right) \qquad \qquad \forall \,(i,k) \in \mathcal{A} \label{cons:dSep}\\

@@

		& x_{ij} \geq 0  \qquad \qquad \forall \,i \in I, j \in J_i \label{cons:dxub}\\
		& t_S = 0. \label{cons:dNonnegt}

@@

	In this formulation, \(t_i\) represents the start time of activity \(i \in I\). We aim to minimize the project span, which is the start time of the terminal activity, \(t_T\). Constraint~\eqref{cons:dSep} guarantees the precedence relationship: if activity \(i\) precedes activity \(k\), activity \(k\) cannot start until activity \(i\) is finished, since the finish time of activity \(i\) is \(t_i + D_{ik} (1- \sum_{j \in J_i} x_{ij}e_{ij})\). Constraint~\eqref{cons:dBudget} is the budget constraint and constraint~\eqref{cons:dSingleBudget} ensures that no more than one unit of crashing option can be applied to an activity. Constraint~\eqref{cons:dNonnegt} enforces nonnegativity for the start time of all activities since activity \(S\) precedes every other activity.\\
	% In constraint~\eqref{cons:dxub}, we can put an artificial upper bound \(1\) on all crashing decisions \(x_{ij}\) by taking proper values of \(e_{ij}\). \\

@@

	When the program evaluation and review technique (PERT) \cite{malcolm1959application} was introduced, activity durations were modeled as independent beta random variables, and the project duration approximated by a normal distribution. Extensions that allow for more general assumptions followed \cite{Elmaghraby77}, and Monte Carlo simulation plays a role in estimating the expected project span, which is difficult to express analytically \cite{burt1971conditional,van1963letter}.
	%The deterministic model can be extended to incorporate uncertainty of activity duration, starting with the program evaluation and review technique (PERT) \cite{malcolm1959application}. The PERT model assumes that activities are independently distributed with beta distributions, and the expected project span can be approximated by a normal distribution. However, this assumption may not be accurate for a general case~\cite{Elmaghraby77}. Therefore, assuming the activity duration follow a given probability distribution, Monte Carlo simulation methods can be used to estimate the expected project span, since it is is difficult to express the expected project span analytically \cite{burt1971conditional,van1963letter}. 
	Heuristics and simulation-based algorithms have been developed to approximately solve the stochastic project crashing problem \cite{aghaie2009ant, bowman1994stochastic, ke2014genetic, kim2007heuristic}. Another approach to handle the uncertainty is robust optimization, in which the objective is to minimize the worst case project span within a specified uncertainty set. While affinely adaptive recourse decisions are computationally tractable as linear, or second-order cone, programs, this restriction may lead to suboptimal solutions \cite{chen2008linear, cohen2007stochastic}. However, once recourse decisions can take general form, the robust model is only tractable with hypercube uncertainty sets, which is a trivial case because the worst-case scenario in the uncertainty set is always the point where every activity takes its upper bound \cite{wiesemann2012robust}. Ahipasaoglu et al. \cite{ahipasaoglu2016distributionally} propose a distributionally robust optimization scheme applied to a PERT network, which reformulates the problem as a semidefinite program or a copositive program, depending on the description of uncertainty. The project crashing optimization problem finds wide applications in project management~\cite{demeulemeester2006project, jaselskis1991allocation,tonchia2018industrial}, machine scheduling~\cite{blazewicz1983scheduling,hall1996machine}, health services scheduling~\cite{cardoen2010operating}, chemical processes~\cite{li2008process}, and digital circuit sizing~\cite{kim2007heuristic}.\\

@@

	In this paper, we propose using the concept of stochastic disruptions to model the uncertainty, which is different from the setting of a stochastic program or a robust optimization problem. A stochastic disruption is an event that may occur at any time in the time horizon and change the system parameters significantly. A few authors apply this idea when the disruption can only occur in a set of specified time periods. Yu and Qi \cite{yu2004disruptionmgt} introduce scenario-based optimization models for airline scheduling. Morton et al. \cite{morton2009sealift} introduce a sealift scheduling problem under a finite number of stochastic disruptions within a stochastic programming structure. This structure ``falls between standard two-stage and multi-stage stochastic programs for a multi-period problem" and reduces the size of the problem to a quadratic growth in the number of time periods. Our setting inherits the philosophy of \cite{morton2009sealift} but enhances the model by allowing a continuous disruption time instead of a prespecified set of fixed time periods. \\

@@

	We first formally describe the crashing optimization problem under stochastic disruptions. Given a limited number of disruption scenarios, the problem can be formulated as a stochastic mixed integer program and we present the extensive formulation in Section~\ref{sec:formulation}. If we assume a continuous distribution for the disruption time and magnitude, a sample average approximation (SAA) can be used to create a finite set of scenarios and approximate the original problem by a finite-sized optimization problem. We present some important properties of this problem using a serial activity network as a special case in Section~\ref{sec:examples}. The large scale and the discrete non-convex nature of the SAA problem's extended formulation indicate that it may not be solved efficiently to desirable tolerance level. In Section~\ref{sec:decomposition}, a branch-and-cut method based on Benders decomposition is developed to solve the crashing optimization problem with a disruption. We show such a decomposition method solves the integer program to the exactness within a finite number of iterations. The experiment results are presented in Section~\ref{sec:results}, including the relationship between the solution quality and the sample size, the comparison between the quality of our solution and the deterministic/semi-stochastic alternative solutions, and the computational performance of the decomposition method presented in Section~\ref{sec:decomposition}. We conclude our paper with remarks on potential extensions of this model in Section~\ref{sec:conclusions}.

@@

	For a project crashing problem under stochastic disruptions, we assume at most one stochastic disruption can occur at a random time in the project span, since for the applications we are interested in (natural disasters, terrorist attacks, major market crashes, etc.), it is unlikely for two disruptions to occur in entire the time horizon. For example, suppose we manage a construction project, and we aim to plan against the potential hazard caused by earthquakes. It is unlikely for two major earthquakes to hit the same area twice within a year (aftershocks can be considered part of a major earthquake.) It is also reasonable to assume for each activity \(i \in I\), the crashing decision needs to be made at the start of that activity, and cannot be changed afterward, especially when contracts are involved in commitment of resources~\cite{oberlender1993project}. The disruption does not affect the started activities but changes the length of those that have not started according to a random distribution. It is usually hard to compute the recourse function directly when parameters are distributed according to continuous distributions, and therefore SAA is proposed to approximate the problem by a large-scale deterministic optimization problem with a finite scenario set \cite{kim2015guide,shapiro2009lectures}. In this paper, we assume there is a discrete finite set of scenarios indexed by \(\omega \in \Omega\). For each scenario \(\omega\), the random realization of parameters, \(\xi^\omega\), consists of the timing of disruption, \(H\), and the changed duration of each activity, \(d_i, \forall i \in I\). \\

@@

	If there is at most one disruption, we can model the problem as a two-stage stochastic mixed integer program. The definition of our first stage is different from the traditional stochastic program setting. Here the first stage contains decisions through completion of the project if no disruption occurs and the second stage characterizes the decisions for each realization of the disruption. The first stage decision variables are carried out until the disruption if it ever occurs, and after the disruption, the scenario-specific recourse decisions are executed.\\

@@

		\(t_{i}\) & \(\qquad\) & nominal start time of activity \(i\), \(i \in I\);&\\

@@

		\(t_{i}^\omega\) & \(\qquad\) & start time of activity \(i\) under scenario \(\omega\), \(i \in I, \omega \in \Omega\);&\\

@@

\section{NP-Hardness} \label{sec:nphard}
	In this section, we prove that the project crashing optimization problem under a stochastic disruption is NP-hard. The proof process is largely based on the proof of De et al.~1997 for the discrete time-cost tradeoff (DTCT) problem. In De et al.~\cite{de1997complexity}, the authors show that an exactly-one-in-three 3SAT (EOIT\_3SAT) problem can be transformed to an instance of the DTCT decision problem in polynomial time. Since EOIT\_3SAT is NP-complete (see \cite{Garey1979ComputersAI} for the detailed proof), the authors prove that the DTCT decision problem is also NP-complete, and the DTCT optimization problem is NP-hard.\\
	Similar to De et al.~\cite{de1997complexity}, we also first introduce the EOIT\_3SAT problem: let \(U = \{u_1,u_2, \dots, u_n\}\) be a set of variables. A literal can be either \(u\) or \(\bar{u} = \neg u\) for a \(u \in U\). Let \(C = \{c_1, c_2, \dots, c_m\}\) be a set of clauses, each of which is formed by a disjunction of three literals. The EOIT\_3SAT problem asks for whether there is a truth assignment for \(U\) such that each clause in \(C\) has exactly one true literal. \\
	Based on the above instance of EOIT\_3SAT, we formulate an instance of the activity network for the project crashing problem. The activity network consists of three layers of nodes: the first layer contains \(4n\) nodes and represents the truth assignment of each variable; the second layer contains \(3m\) nodes and represents the value of clauses; the third layer contains node \(T\) which represents the end of the project. The first layer has \(n\) components in parallel, each of which corresponds to a variable and contains 4 nodes connected, as shown in Figure~\ref{fig:layer1}. Figure~\ref{fig:layer1} also displays the connection between the first layer and the third layer. We set the scenario set as a singleton \(\Omega = \{1\}\) and set the disruption time as \(H^1 = 0\) with probability \(p^1 = 1\). We set parameter values associated with the arcs in Figure~\ref{fig:layer1} as:
	& D_{u_{j1},u_{j2}} = 0 \qquad \quad d^1_{u_{j1},u_{j2}} = 0  \\
	& D_{u_{j1},u_{j4}} = 2 \qquad \quad d^1_{u_{j1},u_{j4}} = -1  \\
	& D_{u_{j2},u_{j3}} = 1 \qquad \quad d^1_{u_{j2},u_{j3}} = 1\\
	& D_{u_{j4},T} = 0 \qquad \quad \;\; d^1_{u_{j4},T} = 0.
	We do not allow any activities within the first layer to be crashed, i.e., \(J_{u_{jk}} = \emptyset\), for all \(j = 1,2,\dots, n,\ k = 1,2,3,4\). With this setup, the length of the critical path for the \(j\)-th component is always \(2\) for all \(j = 1,2,\dots,n\), as it is optimal to have activity \(u_{j1}\) start at time \(0\). However, whether the critical path traverses activities \(u_{j2}\) and \(u_{j3}\) (top path) or the activity \(u_{j4}\) (bottom path) depends on the value of \(G^1_{u_{j1}}\). If \(G^1_{u_{j1}} = 1\), the \(G\) variables will take value \(1\) for all subsequent nodes. The top path yields a length of \(D_{u_{j1},u_{j2}} + d^1_{u_{j1},u_{j2}} + D_{u_{j2},u_{j3}} +d^1_{u_{j2},u_{j3}} = 0 + 0 + 1 + 1 = 2\), while the bottom path yields a length of \(D_{u_{j1},u_{j4}} + d^1_{u_{j1},u_{j4}} = 2 - 1 = 1\), which makes the top path critical. If \(G^1_{u_{j1}} = 0\), it is optimal to have activity \(u_{j2}\) start before the disruption as well, i.e., \(G^1_{u_{j2}} = 0\). The top path yields a length of \(D_{u_{j1},u_{j4} - 2} + D_{u_{j2},u_{j3}} = 0 + 1 = 1\), while the bottom path yields a length of \(D_{u_{j1},u_{j4}} = 2\), which makes the bottom path critical. We can consider the value of \(G_{u_{j1}}\) as the truth assignment of variable \(u_j\). If the variable \(u_j\) is TRUE, the critical path is the top path; if it is FALSE, the critical path is the bottom path.
		\includegraphics[width=0.8\textwidth]{layer1}
		\caption{A component corresponding to variable \(u_j,\ j = 1,2,\dots, n\) in the first layer of the constructed activity network for EOIT\_3SAT.}
		\label{fig:layer1}
	\noi The arcs from activities \(u_{j3}\) and \(u_{j4}\) in Figure~\ref{fig:layer1} point to activities in the second layer. Next, we construct the second layer of activities and the arcs connecting the first layer and the second layer, in a similar manner as Figure 2(b) in~\cite{de1997complexity}. To best illustrate the arc construction, we show an example of a clause, \(c_i = u_j \vee u_k \vee \bar{u}_{\ell}\), with two original variables, \(u_j\) and \(u_k\), and one negative complement, \(\bar{u}_{\ell}\). We can consider the activity \(c_{ip}\) corresponding to the truth assignment of the variable \(u_p\), for \(p \in \{j,k,\ell\}\). For the original variables \(u_p\), \(p \in \{j,k\}\), we connect activity \(u_{p4}\) with activity \(c_{ip}\), and activity \(u_{p3}\) with activities \(c_{iq}\), where \(q \in \{j,k,\ell\}, q \neq p\). For the negative complement variable \(u_\ell\), we connect activity \(u_{\ell 3}\) with activity \(c_{i \ell}\), and activity \(u_\ell 4\) with activities \(c_{iq},\ q \in \{j,k\}\). This example by no means limits the clauses to two variables and one negative complement but aims to showcase the procedure of constructing the network components corresponding to a variable and the negative complement of a variable. From now on we refer to the activities representing variables as \(u\)-activities and those representing clauses as \(c\)-activities.\\
	We let both the nominal duration \(D\) and the disrupted effect \(d\) for the arcs between \(u\)-activities and \(c\)-activities be \(0\). For the arcs between \(c\)-activities and \(T\), we set:
	%The parameters of arcs in Figure~\ref{fig:layer2} are set as:
	%	& D_{u_{p3},c_{iq}} = 0 \qquad \quad d^1_{u_{p3},c_{iq}} = 0 \quad \forall p \in \{j,k\}, q \in \{j,k,\ell\}, p \neq q\\
	%	& D_{u_{p4},c_{ip}} = 0 \qquad \quad d^1_{u_{p4},c_{ip}} = 0 \quad \forall p \in \{j,k\}\\
	%	& D_{u_{\ell 3},c_{i\ell}} = 0 \qquad \quad d^1_{u_{\ell 3},c_{i \ell}} = 0\\
	%	& D_{u_{\ell 4},c_{iq}} = 0 \qquad \quad d^1_{u_{\ell 4},c_{iq}} = 0 \quad \forall q \in \{j,k\}\\
		& D_{c_{iq},T} = 1 \qquad \quad \;\; d^1_{c_{iq},T} = 0 \;\;\;\quad \forall q \in \{j,k,\ell\}.
	Unlike the \(u\)-activities, each \(c\)-activity can be crashed with a single option with effectiveness \(e_{c_{ip}1} = 1\), \(\forall p \in \{j,k,\ell\}\). The total budget, \(B\), is then set to \(B = 2m\), where \(m\) is the total number of clauses.
		\includegraphics[width=0.5\textwidth]{layer2}
		\caption{The \(i\)-th clause, \(u_{j} \vee u_{k} \vee \bar{u}_{\ell}\), in the second layer of the constructed activity network for EOIT\_3SAT with the arcs connecting it with the first and the third layer.}
		\label{fig:layer2}
	\noi For simplicity, the reasoning behind this construction is stated using the example in Figure~\ref{fig:layer2}: for variable \(u_j\), if \(G^1_{u_{j1}} = 1\), the activity \(u_{j4}\) can start at time \(1\), while the earliest time the activity \(u_{j3}\) can start is \(2\). If \(G^1_{u_{j1}} = 0\), the activity \(u_{j3}\) can start at time \(1\), while the earliest time the activity \(u_{j4}\) can start is \(2\). This is the same for variable \(u_k\) but directly opposite for variable \(u_\ell\). The truth assignments indicate which path is critical, which is similar in De et al.~\cite{de1997complexity}, where the truth assignments indicate which crashing option to select for variables nodes. \\
	%Without loss of generality, suppose \(u_j\) is the only TRUE assignment, then activity \(c_{ij}\) can start as early as time \(1\), while the other two \(c\)-activities can only start after time \(2\). 
	We can prove the following lemmas. Without loss of generality, a clause can be made of without negative complement of variables. For clarity, Figure~\ref{fig:layer2} equivalent for a clause without negative complement is displayed as Figure~\ref{fig:layer2nonneg}, which is used to illustrate the proofs for Lemma~\ref{lemma:onlyOne} and Lemma~\ref{lemma:iff}.
		\includegraphics[width=0.5\textwidth]{layer2nonneg}
		\caption{The \(i\)-th clause, \(u_{j} \vee u_{k} \vee u_{\ell}\), in the second layer of the constructed activity network for EOIT\_3SAT with the arcs connecting it with the first and the third layer.}
		\label{fig:layer2nonneg}
	\begin{lemma} \label{lemma:onlyOne}
		For any clause \(c_i = u_j \vee u_k \vee u_\ell,\ i = 1,2,\dots,m\), there is at most one activity \(c_{iq^*},\ q^* \in \{j,k,\ell\}\) that has the earliest start time as \(1\), and the start time for \(c_{iq}\) is \(2\), for \(q \in \{j,k,\ell\}, q \neq q^*\).
	\end{lemma}
	\begin{proof}[Proof of Lemma \ref{lemma:onlyOne}]
		Without loss of generality, suppose both \(c_{ij}\) and \(c_{ik}\) can start at time \(1\). This means that both \(u_{j3}\) and \(u_{j4}\) needs to start at time \(1\), which is impossible because no matter what value \(G^1_{u_{j1}}\) takes, at least one activity in \(\{u_{j3},u_{j4}\}\) has to start after time \(2\).
	\begin{lemma} \label{lemma:iff}
		For any clause \(c_i = u_j \vee u_k \vee u_\ell,\ i = 1,2,\dots,m\), one activity \(c_{iq^*},\ q^* \in \{j,k,\ell\}\) has the earliest start time as \(1\) if and only if variable \(\{u_{q^*}\}\) is the only variable in the clause to which the truth assignment is TRUE.
	\end{lemma}
	\begin{proof}[Proof of Lemma \ref{lemma:iff}]
		Without loss of generality, we assume \(q^* = j\).\\
		\(\implies\):  Suppose \(u_j\) is FALSE, then activity \(u_{j4}\) can only start after time \(2\), which will lead to the contradiction that \(c_{ij}\) can start as early as time \(1\). Suppose either \(u_k\) or \(u_\ell\) is TRUE in addition to \(u_j\). Since there is going to be an arc from either activity \(u_{k3}\) or \(u_{\ell 3}\) to activity \(c_{ij}\), and either \(u_{k3}\) or \(u_{\ell_3}\) cannot start before time \(2\), it leads to a contradiction that \(c_{ij}\) can start as early as time \(1\). Therefore, if activity \(c_{ij}\) can start at time \(1\), then \(u_j\) is the only variable which is TRUE.\\
		\(\impliedby\): if only \(u_j\) is TRUE, then \(u_{j4}, u_{k3}\) and \(u_{\ell 3}\) can all start as early as time \(1\), which means that the earliest start time of \(c_{ij}\) is \(1\).
	%For a clause \(i\) that contains variable \(j\), activity \(u_{j4}\) connects to \(c_{ij}\) and activity \(u_{j3}\) connects to the \(c\)-activities which do not correspond to variable \(u_j\). This means that if variable \(u_j\) is assigned a TRUE value, activity \(c_{ij}\) can start as soon as time \(1\) and the other two \(c\)-activities can start no earlier than time \(2\). 
	\noi A direct result from Lemma~\ref{lemma:iff} shows if there exists a truth assignment that meets the requirement of EOIT\_3SAT, there are at most \(2m\) number of \(c\)-activities (2 per clause) that has to start after time \(2\). Since we have a total of \(B = 2m\) units of resource, we can crash all of those \(c\)-activities to achieve a project length of \(2\). This means that we can transform an EOIT\_3SAT problem to a project crashing decision problem under a stochastic disruption using the activity network construction process above. If we can find a feasible EOIT\_3SAT assignment, it is possible to obtain a project length less than or equal to 2. Next, we show the formal NP-hardness proof of the project crashing problem with a stochastic disruption. We first formally define the project crashing decision problem under a stochastic disruption:
		A project crashing decision problem under a stochastic disruption is: if we solve is model~\ref{prob:extensive}, there a feasible \(\{t,x,G\}\) such that the optimal expected project length is smaller than or equal to \(\mathcal{T}\).
	\begin{theorem}\label{thm:npcomplete}
		For an instance of the EOIT\_3SAT problem, we construct the corresponding project crashing decision problem under a stochastic disruption: let \(\Omega = \{1\}\), \(H^1 = 0\), \(\mathcal{T} = 2\), and the precedence relationships, the crashing options and budget are constructed by the process shown in the Section~\ref{sec:nphard}. This decision problem has a solution if and only if the given instance of EOIT\_3SAT problem has a solution.
	\begin{proof}[Proof of Theorem~\ref{thm:npcomplete}]
		(polynomial transformation) 
		%We first show that any instance of the project crashing decision problem is constructed in time that is polynomial in the length of the given instance of EOIT\_3SAT, and has a length that is polynomially related to the length of that instance. 
		Suppose the EOIT\_SAT problem has \(n\) variables and \(m\) clauses, the constructed activity network for the project crashing problem has \(4n + 3m + 1\) activities and \(5n + 12m\) arcs, which are both polynomial in the size of the original EOIT\_SAT problem. And all numbers in the project crashing decision problem are smaller than a constant, \(2\).\\
		(\(\implies\)) Suppose the EOIT\_SAT instance has a solution. We have shown that we can start every activity as early as possible, and spend the budget to crash all \(c\)-activities which correspond to the literals with FALSE assignment and have the start time at \(2\). \\
		(\(\impliedby\)) Suppose the project crashing decision problem has a solution \((\hat{t},\hat{x},\hat{G})\) where \(t_T = 2\). By Lemma~\ref{lemma:onlyOne} we know that for each clause there are at most one \(c\)-activity that can start at time \(1\), which means in total there must be at least \(2m\) number of \(c\)-activities with a start time later than \(2\). Since the crashing budget is exactly \(2m\), if there are more than \(2m\) \(c\)-activities that starts after time \(2\), the project length will be larger than 2, which means that there are exactly \(2m\) number of \(c\)-activities starting at time \(2\). This is equivalent to that exactly one \(c\)-activity in each of the \(m\) clauses starts at time \(1\). From Lemma~\ref{lemma:iff} we can directly obtain the result that for each clause there is exactly one variable to which the truth assignment is TRUE, which is essentially a solution to the EOIT\_3SAT instance.\\
		EOIT\_3SAT is a NP-complete problem~\cite{Garey1979ComputersAI}. The project crashing decision problem under a stochastic disruption in NP, since we can check in \(O(n^2)\) to check whether a given input is feasible and has \(t_T \leq \mathcal{T}\). This complete the proof that the project crashing decision problem under a stochastic disruption is NP-complete.
	\begin{theorem} \label{thm:nphard}
		The project crashing optimization problem under a stochastic disruption is NP-hard.
	\begin{proof}[Proof of Theorem~\ref{thm:nphard}]
		From Theorem~\ref{thm:npcomplete} states that the project crashing decision problem under a stochastic disruption is NP-complete. This implies the project crashing optimization problem under a stochastic disruption, formulated as model~\ref{prob:extensive}, is NP-hard.
%	\noi For the network \(G(I,\mathcal{A})\) constructed by the process above, we introduce the decision problem: suppose the disruption time takes a constant value \(0\), is there a set of start time \(t_i,\ i \in I\), a set of crashing plan \(x_{ij},\ i \in I, j \in J_i\) and a set of indicators \(G_i\) showing whether activity \(i\) starts before the disruption or after.

@@

	In this section, we show examples of serial activity networks to illustrate the nature of the project crashing problem under stochastic disruption, which is different from the deterministic problem. In the deterministic project crashing problem, all activities on the critical path should start as soon as possible. However, with a stochastic disruption, it might be optimal to delay the start of some activities. Therefore, it is necessary to have decision variable \(t_i, \forall i \in I\) in model~\eqref{prob:extensive} in addition to the crashing decision variable \(x_{ij}, \forall i \in I, j \in J_i\). \\

@@

	Because it is possible for an optimal crashing plan to contain a delay for some activities, we have to set up decision variables \(t_i,\ \forall i \in I\), as the start time of each activity, rather than assume that each activity starts as soon as all of its predecessors are finished.  We summarize the reasons for such delay include:

@@

			\item Some activities might have a shorter length after the disruption. It is beneficial to wait a short period of time to capture this advantage.
